---
title: "Assignment 7 - Visualization - Some aspects of ggplot2"
format: html
editor: visual
---

***How to do it?***:

- Open the Rmarkdown file of this assignment ([link](assignment7.qmd)) in Rstudio.

- Right under each **question**, insert a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question.

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.

- Publish the html file to your Githiub Page.

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

------------------------------------------------------------------------

Use the data of your own. Produce the following types of plots and comment on each plot. Plots should be meaningful. If you use the data we used in class, make sure the plots are not the same as the ones in the slides. All plots should have title, caption, appropriate labels on x and y-axis.

------------------------------------------------------------------------

1.  Make a plot using `facet_wrap` by one categorical variable (`facet_wrap(~variable1)`)

    ```{r}
    library(tidyverse)
    library(knitr)
    library(ggplot2)
    df <- read_csv("C://Users//student//Documents//GitHub//Math421_Assignments//titanic.csv")
    df <- df %>% mutate(Survived = as.character(Survived), Pclass = as.character(Pclass))
    df %>% ggplot()+ geom_point(mapping=aes(x=Age, y=Fare))+facet_wrap(~Sex)+
    labs(title = "Titanic Fare by Passenger Sex", x = "Age of Passenger", y = "Cost of Fare", fill = "Survived", caption = "There is no correlation between passenger age and fare cost") 
    ```

2.  Make a plot using `facet_wrap` by two categorical variables (`facet_wrap(~variable1+variable2)`)

    ```{r}
    df %>% ggplot()+geom_point(mapping=aes(x=Age, y=Fare))+facet_wrap(~Pclass+Survived)+
      labs(title = "Titanic Fare by Passenger sex and class", x = "Age of Passenger", y = "Cost of Fare", fill = "Survived", caption = "First class survivors had the highest average fare") 
    ```

3.  Make a plot using `position ='dodge'`.

    ```{r}
    df %>% ggplot()+geom_bar(mapping=aes(x=Pclass, fill=Survived), position = 'dodge')+
    labs(x='Passenger Class', y='Number of Passengers', fill='Survived', title='Survival Counts by Passenger Class', caption='Side-by-side bars compare survival outcomes.')
    ```

4.  Make a plot using `position ='fill'`.

    ```{r}
    df %>% ggplot()+geom_bar(mapping=aes(x=Pclass, fill=Survived), position = 'fill')+
      labs(x='Passenger class', y='Survival rate', fill='Survived', title='Survival Proportions by Passenger Class', caption='Each bar represents the survival rate of all passengers in that class.')
    ```

5.  Make a plot using `geom_col`.

    ```{r}
    df %>% group_by(Sex) %>% summarise(average_fare=mean(Fare, na.rm=TRUE)) %>% ggplot()+ geom_col(mapping=aes(x=Sex, y=average_fare))+
      labs(x='Sex', y='Average Fare', title='Average Fare by Sex',
    caption='The average fare for females was much higher than for males')
    ```

6.  Change the theme of one of your plots.

    ```{r}
    df %>% group_by(Sex) %>% summarise(average_fare=mean(Fare, na.rm=TRUE)) %>% ggplot()+ geom_col(mapping=aes(x=Sex, y=average_fare))+ theme_linedraw()+
      labs(x='Sex', y='Average Fare', title='Average Fare by Sex', caption='The average fare for females was much higher than for males')
    ```

7.  Make a plot using combo filter + ggplot

    ```{r}
    df %>% filter(Age>=20, Age<=40) %>% ggplot()+geom_bar(mapping=aes(x=Pclass, fill=Survived))+
      labs(x='Passenger Class', y='Number of Passengers', fill='Survived', title='Survival of Passengers Aged 20 to 40', caption='First class had the highest survival rate within the age range')
    ```

8.  Make a plot using combo group_by + summarise + geom_col

    ```{r}
    df %>% group_by(Survived, Sex) %>% summarise(mean_fare=mean(Fare, na.rm=TRUE)) %>% ggplot()+geom_col(aes(x=Sex, y=mean_fare, fill=Survived), position = 'dodge')+
      labs(x='Gender', y='Average Fare', fill='Survived', title='Average Fare by Gender and Survival', caption='A higher fare correlates to a higher survival rate regardless of gender')
    ```

9.  Save a plot and put a link of the plot to your Github Webpage. To link a plot `abc.png` you just need to insert the following on the README.md

```{r, eval=FALSE}
gg <- df %>% group_by(Sex) %>% summarise(average_fare=mean(Fare, na.rm=TRUE)) %>% ggplot()+ geom_col(mapping=aes(x=Sex, y=average_fare))
ggsave(filename = 'Assignment7Graph.png', plot = gg)
```

10. In this question, we will use the Covid19 vaccination data for the US. The data and data dictionary can be found at the below link:

- [Data](https://bryantstats.github.io/math421/data/vaccinations-by-manufacturer.csv)

```{r}
library(tidyverse)
df2 <- read_csv("C://Users//student//Documents//GitHub//Math421_Assignments//vaccinations-by-manufacturer.csv")

```

`Geom_point` makes scatter plots, which plot two numeric (or Date) variables. A third categorical variable is often brought in for color. For example, we can compare different vaccine types used in the US by

```{r}
df2 %>% filter(location=='United States') %>% ggplot()+geom_point(mapping = aes(x = date, y = total_vaccinations, color = vaccine))
```

Do the follows.

- Use `geom_point` to compare the vaccines used by date in Argentina in 2022.

  ```{r}
  df2 %>% filter(location=='Argentina', date>='2022-01-01', date<'2023-01-01') %>% ggplot()+geom_point(mapping = aes(x = date, y = total_vaccinations, color = vaccine))+
    labs(x='Date', y='Total Vaccinations', color='Vaccine', title='COVID-19 Vaccinations in Argentina, 2022', caption='Vaccination totals by vaccine type.')
  ```

- Change `geom_point` to `geom_line` in the above codes to have the corresponding line plots.

  ```{r}
  df2 %>% filter(location=='Argentina', date>='2022-01-01', date<'2023-01-01') %>% ggplot()+geom_line(mapping = aes(x = date, y = total_vaccinations, color = vaccine))+
    labs(x='Date', y='Total Vaccinations', color='Vaccine', title='Vaccination Trends in Argentina, 2022', caption='Lines show recorded totals by vaccine type.')
  ```

11. (Using the same dataset as in 10.) Sometime we need to do some calculations on the data before plotting. For example, the below codes is to compare the total vaccines (including all types of vaccines) by date between the US and the EU

```{r, eval=FALSE}
# the total vaccines (including all types of vaccines) by date
d1 <- df %>% 
  filter(location %in% c('United States', 'European Union')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE))

# Plot
d1 %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location))
```

Notice that the above codes could be piped in one

```{r, eval=FALSE}
df2 %>% 
  filter(location %in% c('United States', 'European Union')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE)) %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location))
```

Do the follows.

- Use `geom_point` compare the total vaccines (including all types of vaccines) by date between countries/regions in Asia

  ```{r}
  df2 %>% filter(location %in% c('China', 'India', 'Japan', 'Indonesia', 'Thailand', 'Malaysia', 'Singapore', 'Philippines', 'Vietnam', 'South Korea')) %>% group_by(date, location) %>% summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE)) %>% ggplot()+ geom_point(mapping=aes(x=date, y=total_vac2, color=location))+
    labs(x='Date', y='Total Vaccinations', color='Country', title='Vaccination Totals in Selected Asian Countries', caption='Recorded totals by date and country.')
  ```

- Use `geom_point` compare the total vaccines (including all types of vaccines) by date between countries/regions in South America in 2021

  ```{r}
  df2 %>% filter(location %in% c('Argentina', 'Brazil', 'Chile', 'Colombia', 'Peru', 'Uruguay', 'Ecuador', 'Bolivia', 'Paraguay', 'Venezuela', 'Guyana', 'Suriname'), date>='2021-01-01', date<'2022-01-01') %>% group_by(date, location) %>% summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE)) %>% ggplot()+geom_point(mapping=aes(x=date, y=total_vac2, color=location))+
    labs(x='Date', y='Total Vaccinations', color='Country', title='Vaccination Totals in South America, 2021', caption='Recorded totals by date and country.')
  ```
