---
title: "Assignment 3"
format: html
editor: visual
---

***How to do it?***:

-   Open the Quarto file of this assignment ([link](assignment3.qmd)) in Rstudio.

-   Right under each **question**, insert a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question.

-   `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.

-   Publish the html file to your Githiub Page.

***Submission***: Submit the link on Github of the assignment to Canvas.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

------------------------------------------------------------------------

## Problems

\

1.  Create the following data frame

| Rank | Age | Name  |
|:-----|:----|:------|
| 0    | 28  | Tom   |
| 1    | 34  | Jack  |
| 2    | 29  | Steve |
| 3    | 42  | Ricky |

```{r}
df = data.frame(Rank = c(0,1,2,3),
                Age = c(28,34,29,42),
                Name = c('Tom','Jack','Steve','Ricky'))
df
```

2.  Use `read.csv` to import the Covid19 Vaccination data from WHO: [link](https://bryantstats.github.io/math421/data/us-states.csv).

    ```{r}
    df2 <- read.csv("C:\\Users\\student\\Documents\\GitHub\\Math421_Assignments\\us-states.csv")
    head(df2)
    ```

3.  Show the names of the variables in the data

```{r, echo=FALSE}
names(df2)

```

4.  How many columns and rows the data have?

    ```{r}
    nrow(df2)
    ncol(df2)
    ```

5.  How many missing values are there? Show the missing values by columns. What variable has the most number of missing values?

    ```{r}
    sum(is.na(df2))
    colSums(is.na(df2))
    colMeans(is.na(df2))
    ```

6.  What is the class of the `date` column. Change the `date` columns to `date` type using the `as.Date` function. Show the new class of the `date` column.

    ```{r}
    class(df2$date)
    df2$date = as.Date(df2$date)
    class(df2$date)
    ```

7.  Capitalize the names of all the variables

    ```{r}
    names(df2)[1] <- 'Date'
    names(df2)[2] <- 'State'
    names(df2)[3] <- 'Fips'
    names(df2)[4] <- 'Cases'
    names(df2)[5] <- 'Deaths'
    names(df2)
    ```

8.  Find the average number of cases per day. Find the maximum cases a day.

    ```{r}
    mean(df2$Cases, na.rm = TRUE)
    max(df2$Cases, na.rm = TRUE)
    ```

9.  How many states are there in the data?

    ```{r}
    length(unique(df2$State))
    ```

10. Create a new variable `weekdays` to store the weekday for each rows.

    ```{r}
    df2$weekdays <- weekdays(df2$Date)
    head(df2)
    ```

11. Create the categorical variable `death2` variable taking the values as follows

-   `has_death` if there is a death that day
-   `no_death` if there is no death that day

Find the frequency and relative frequency of `no_death` and `has_death`.

```{r}
df2$death2 <- ifelse(df2$Deaths > 0, "has_death", "no_death")
table(df2$death2)
prop.table(table(df2$death2))
```

12. Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable `death`. (Hint: Use the `summary` function)

    ```{r}
    summary(df2$Deaths)
    ```

13. Create the categorical variable `death3` variable taking the values as follows

-   `low_death` if the number of deaths smaller than the 25 percentile (Q1)

-   `mid_death` if the number of deaths from Q1 to Q3

-   `high_death` if the number of deaths greater than Q3

    ```{r}
    Q1 <- quantile(df2$Deaths, 0.25, na.rm = TRUE)
    Q3 <- quantile(df2$Deaths, 0.75, na.rm = TRUE)
    df2$death3 <- ifelse(
      df2$Deaths < Q1, "low_death",
      ifelse(df2$Deaths > Q3, "high_death", "mid_death"))
    table(df2$death3)
    ```

14. Find the average cases in Rhode Island in 2021

    ```{r}
    mean(df2$Cases[df2$State == "Rhode Island" &
                  format(df2$Date, "%Y") == "2021"],
      na.rm = TRUE)
    ```

15. Find the median cases by weekdays in Rhode Island in 2021

    ```{r}
    RI2021 <- df2[df2$State == "Rhode Island" &
      format(df2$Date, "%Y") == "2021",]
    by(RI2021$Cases, RI2021$weekdays, median, na.rm = TRUE)
    ```

16. Compare the median cases in Rhode Island in June, July, August and September in 2021.

    ```{r}
    #%y is year %m is month
    RI2021 <- df2[df2$State == "Rhode Island" &
      format(df2$Date, "%Y") == "2021",]
    RI2021$month <- format(RI2021$Date, "%m")
    by(RI2021$Cases,
      RI2021$month,
      median,
      na.rm = TRUE)
    ```

17. Find your own dataset, import it and implement the following functions on the data

-   head

-   str

-   names

-   mean, min, max

-   table

-   is.na

-   colSums

-   class

-   cor

-   by

-   ifelse

-   case_when

    ```{r}
    titanic <- read.csv("https://bryantstats.github.io/math421/data/titanic.csv")
    head(titanic)
    str(titanic)
    names(titanic)
    mean(titanic$Age, na.rm = TRUE)
    min(titanic$Age, na.rm = TRUE)
    max(titanic$Age, na.rm = TRUE)
    table(titanic$Sex)
    is.na(titanic)
    colSums(is.na(titanic))
    class(titanic$Age)
    cor(titanic$Age, titanic$Fare, use = "complete.obs")
    by(titanic$Fare, titanic$Sex, mean, na.rm = TRUE)
    titanic$AgeGroup <- ifelse(titanic$Age < 18,"Child","Adult")
    library(dplyr)
    titanic$AgeGroup2 <- case_when(titanic$Age < 18 ~ "Child",titanic$Age >= 18 & titanic$Age < 65 ~ "Adult",
    titanic$Age >= 65 ~ "Senior",
    TRUE ~ "Unknown")
    ```

18. In the dataset in #17, practice the follows. You can reuse the code of 17.

-   Create a categorical variable from a continuous variable

-   From one categorical variable, create a new categorical variable with fewer categories

    ```{r}
    titanic$FareGroup <- ifelse(titanic$Fare < 20,"Low",ifelse(titanic$Fare < 50,"Medium","High"))
    table(titanic$FareGroup)
    titanic$ClassGroup <- ifelse(titanic$Pclass == 1,"First","Second_or_Third")
    table(titanic$ClassGroup) 
    ```

If you do not have a data, you can use [the titanic dataset](https://www.kaggle.com/competitions/titanic/overview), which can be downloaded at [this link](https://bryantstats.github.io/math421/data/titanic.csv)
